\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2018-01-29 eg6-letters/NIM/EG6_RTPC_NIM.tex
correct some typos

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis. It is a 6~mm diameter Kapton straw with a 27~\textmu{}m wall 
     of 292~mm length such that its entrance and exit 15~\textmu{}m aluminum windows 
     are placed outside of the detector volume. The detector and the target are 
     placed in the center of the solenoid, 64~cm upstream of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 \textmu{}m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~\textmu{}m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The electric field in this region is perpendicular to the beam and 
      average around 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

A new custom backplane was developed to connect FECs to the Readout Control 
Unit (RCU)~\cite{RCU}, allowing fast (200 MB/s) communication between the 
boards. The RCU board, used to distribute the trigger signal to different FECs 
and read data, was equipped with a fast optical link for data transferring to 
the main event builder. These features, together with the ``block-transfer'' 
readout mode making use of the FECs multi-event buffer, allowed to reach a 
signifantly higher readout-rate compared to the original BoNuS system. During 
the 2009 run, the system was successfully operated with a DAQ rate of 3.1 kHz 
and a live time of 70~\%, for a luminosity of about 
$10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 6.064~GeV, to be compared 
with the 500 Hz obtained with the first BoNuS detector at similar run 
conditions.

@@

The results after calibration are shown in Figure~\ref{fig:dedx_p_exp_2nd}, where energy 
loss of particles is plotted against momentum over charge ratio. One can 
clearly see there the band for $^4$He in its expected position\footnote{The lower peak 
   in the left module of the RTPC comes from an unidentified problem in this half 
   of the detector that concerns 7\% of all the elastic events. 
   Our best guess is that these events are linked with a high voltage supply 
   issue, that would sometimes bring the GEM gain down. These events pass all 
   the elastic requirements and we found nothing that differenciates them 
   from other events but their low recorded ADCs. In particular, 94\% of pads
   are involved with these tracks such that we are sure they do not come from 
   a miscalibration of a part of the detector. Finally, we decided to discard
   these events for the calibration procedures.}.

>>>>

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis. It is a 6~mm diameter Kapton straw with a 27~\textmu{}m wall 
     of 292~mm length such that its entrance and exit 15~\textmu{}m aluminum windows 
     are placed outside of the detector volume. The detector and the target are 
     placed in the center of the solenoid, 64~cm upstream of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 \textmu{}m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~\textmu{}m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The electric field in this region is perpendicular to the beam and 
      averages around 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

A new custom backplane was developed to connect FECs to the Readout Control 
Unit (RCU)~\cite{RCU}, allowing fast (200 MB/s) communication between the 
boards. The RCU board, used to distribute the trigger signal to different FECs 
and read data, was equipped with a fast optical link for data transferring to 
the main event builder. These features, together with the ``block-transfer'' 
readout mode making use of the FECs multi-event buffer, allowed to reach a 
significantly higher readout-rate compared to the original BoNuS system. During 
the 2009 run, the system was successfully operated with a DAQ rate of 3.1 kHz 
and a live time of 70~\%, for a luminosity of about 
$10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 6.064~GeV, to be compared 
with the 500 Hz obtained with the first BoNuS detector at similar run 
conditions.

@@

The results after calibration are shown in Figure~\ref{fig:dedx_p_exp_2nd}, where energy 
loss of particles is plotted against momentum over charge ratio. One can 
clearly see there the band for $^4$He in its expected position\footnote{The lower peak 
   in the left module of the RTPC comes from an unidentified problem in this half 
   of the detector that concerns 7\% of all the elastic events. 
   Our best guess is that these events are linked with a high voltage supply 
   issue, that would sometimes bring the GEM gain down. These events pass all 
   the elastic requirements and we found nothing that differentiates them from 
   other events but their low recorded ADCs. In particular, 94\% of pads
   are involved with these tracks such that we are sure they do not come from 
   a miscalibration of a part of the detector. Finally, we decided to discard
   these events for the calibration procedures.}.
